Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Flow

### 2022-08-01 paper/paper.tex

Here, we consider Taylor-Couette (TC) flows at moderate inner and outer Reynolds numbers, $|\Reyn_{i,o}| < 3500$. This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$. 
%GQL has been show to outperform QL models in numerous highly turbulent flows, driven by $\beta$-effects on rotating spheres, rotating convection, rotating plane Couette flow (pCf, here after). 
One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space. In the linearly stable regions, transient growth has been observed in TC flow\cite{2002PhFl...14.3475H, 2002PhFl...14.1655M}. 
We can always write the Navier-Stokes equations in schematic form as
\begin{equation}
    \label{eqn:NS-schematic}
    \frac{\partial \mathbf{q}}{\partial t} + \Lop \mathbf{q} = \mathcal{N}(\mathbf{q}),
\end{equation}
where $\Lop$ is a linear operator and $\mathcal{N}$ is a non-linear function of the state vector $\mathbf{q} = \left[ \mathbf{u}  p\right]^T$, where $\mathbf{u}$ is the velocity field and $p$ the pressure.
As noted by \cite{2002PhFl...14.3475H}, the linear operator $\Lop$ for TC flow is a function of its control parameters.
That is, its commutator with its adjoint
\begin{equation}
    [\Lop^\dagger, \Lop] = f(\mu, \eta, \Reyn_i, \Gamma)
\end{equation}
can be zero or non-zero. 
This provides a very useful arena to test the validity of GQL in a system that admits a varying degree of non-normality.

As a first experiment to this end, we consider the \emph{spiral turbulence} flow regime. The classic paper by Coles \cite{1965JFM....21..385C} was the first to note a hysteretic transition to spiral turbulence when the cylinders were counter-rotating: an experiment at fixed $\Reyn_o$ will have spiral turbulence at a certain $\Reyn_i$ if $\Reyn_i$ is reached by \emph{decreasing} from a higher value, but will not have spiral turbulence at the same $\Reyn_i$ if it is reached by \emph{increasing} from a lower value. More recent work has identified both supercritical and subcritical paths to spiral turbulence\cite{2009PhRvE..80d6315M}. In that work, the authors follow spiral turbulence along a decreasing $\Reyn_i$ path and show that the state is maintained even below the linear stability boundary when $\Reyn_o = -3000$ but it relaminarizes below the stability boundary when $\Reyn_o = -1200$.
Using \cite{2009PhRvE..80d6315M} as a guide, we use direct numerical simulations (DNS) at $\Reyn_o = -3398$ and decrease $\Reyn_i$ into the subcritical region. We then run a suite of GQL simulations in both the super- and subcritical regions.
%Using these prior works as a guide, we consider three cases
%\begin{enumerate}
%    \item Inner cylinder rotating only $\Reyn_o = 0$
%    \item Counter rotating cylinders with moderate outer cylinder rotation $\Reyn_o = -1339$
%    \item Counter rotating cylinders with high outer cylinder rotation $\Reyn_o = -3389$
%\end{enumerate}
%Case 1 has a purely normal linear operator, while the latter two feature moderate and significant non-normality respectively.

@@

We use the pseudo-spectral Dedalus framework (version 2) \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations. We discretize the three velocity components and pressure onto a basis of Chebyshev polynomials in the radial direction, while $\theta$ and $z$ use Fourier bases.
To evolve the solution forward in time, we use the implicit-explicit Runga-Kutta RK443 time stepper; linear terms are implicitly timestepped while non-linear terms are explicit.
Our initial value solver has been extensively tested for TC flow against the linear and non-linear test cases in \cite{1984JFM...146...45M, 1984JFM...146...65M} including linear growth rates, non-linear wave speeds, and torque measurements; our spiral turbulence DNS results also agree with those in \cite{2009PhRvE..80d6315M,2009PhRvE..80f7301D,2011JFM...668..150D}. 
Resolutions and geometric parameters for all runs are given in table~\ref{tab:simulations}.
We have confirmed our simulations are resolved by ensuring that the highest wavenumber components have spectral power at least $10^5$ smaller than the most energetic scale.

Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
\section{Linear Stability}
\label{sec:linear}
In order to determine the linear stability bound, we find the minimum $\Reyn_i$ for a given $\Reyn_o$ by solving the eigenvalue problem for equations~\ref{eqn:NS}. We note that for all $\Reyn_o$ plotted in figure~\ref{fig:LSB}, most unstable mode is never axisymmetric. At $\Reyn_o=-3398$, where our non-linear calculations are run, the most unstable mode has $m = 5$ at onset, and 
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{../figs/reo_rei_lsb.pdf}
    \caption{A summary of spiral turbulence runs in the $\Reyn_o-\Reyn_i$ plane showing linear stability boundary. Solid circles shows points where we have conducted GQL analysis; crosses represent DNS steps along the path to subcritical behaviour. }
    \label{fig:LSB}
\end{figure}

@@

\section{Results}
\label{sec:nonlinear}
%The ($\eta$, $\Gamma$) choice is motivated by \cite{1984JFM...146...45M,1984JFM...146...65M}, particularly section 4 of \cite{1984JFM...146...65M}.

%First, consider a set of runs with $\mu = 0$. This is a standard, linearly unstable case. Here, we choose three $\Reyn$ values, one in the WVF flow regime $\Reyn =750$, one in the MWF $\Reyn = 1500$, and one in the TTV regime $\Reyn = 2000$. We perform two sets of runs, each varying $\Lambda = {0, 1, 2, 6, 10, 20}$ for both the $z$ and $\theta$ directions. The first set of runs is a cold start from identical initial conditions. This allows us to understand the ability of GQL to find certain flow states. The second set of runs will 
We first consider the supercritical case with $\Reyn_i = 900$. We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.  Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$. The first important thing to note is that none of these low-order models are able to maintain the spiral structure. Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda$. Focusing on $\Lambda=1$, we see that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns: the interpenetrating spirals (IPS) and intermittency regimes (INT) originally identified in experiments \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../figs/urms_tz_rei_700_reo-3398_000278.png}
    \caption{Slice of fluctuating rms velocity in the middle of the radial domain at $\Reyn_i = 700$, $\Reyn_o = -3398$. This solution is well below the linear stability threshold at }
    \label{fig:urms_tz_rei700}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figs/rei900_snapshots.pdf}
    \caption{RMS velocity perturbations in the $z-\theta$ plane at $r=R_{mid}$ for $\Reyn_i=900, \Reyn_0=-3398$. The upper left panel shows DNS, the upper right panel shows QL, and the remaining four are GQL with $\Lambda = 1,3,5,10$. Even at $\Lambda = 10$, spiral turbulence is not apparent.}
    \label{fig:rei900_snapshots}
\end{figure}

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. On the left,  $\Reyn_i = 900$ is above the linear stability boundary, while the right shows $\Reyn_i = 700$, below the stability boundary. Starting from a saturated spiral turbulence state at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10$. For $\Reyn_i = 700$, we also use $\Lambda = 20, 30$ At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. }
    \label{fig:ke_vs_t_rei900}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figs}
    \caption{Caption}
    \label{fig:my_label}
\end{figure}

@@

\begin{table}[!h]
\caption{Simulations}
\label{tab:simulations}
\begin{tabular}{rlllllllll}
\hline
Run & $\Reyn_i$ & $\Reyn_o$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$, $\Lambda_z$ & Notes\\
\hline
%A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
%B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
%C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
%\hline
A & 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10 & Supercritical SPT\\
B & 700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10, 20, 30 & Subcritical SPT\\
%\hline
%F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
%G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
\end{tabular}
\vspace*{-4pt}
\end{table}%%%End of the table

%\section{Resolution}
%Might compare poorly resolved DNS to low order GQL as done in\cite{2017JFM...810..412T}. Note that in poorly resolved TC DNS, one gets spurious spiral modes.

%\section{Conjugate Stable Modes}
%Conjugate stable modes? Typically, one considers the effect of the inviscid stable mode rather than the viscously damped mode. However, in the case of TC flow, the background flow profile could not be maintained by the cylinder walls in the limit $\nu \to 0$. However, because circular Couette flow is independent of $\Reyn$, probably the limit is OK? Can think about this. Obviously Rayleigh criterion applies in inviscid case, and this can be easily checked for any angular momentum profile $r^2 \Omega(r)$, so there is some merit to thinking about ideal case even in TC flow. Need to think more about this/run it by Adrian.

%The output for figure is:\vspace*{-7pt}

%\begin{figure}[!h]
%\centering\includegraphics[width=2.5in]{xxxxxx.eps}
%%% where xxxxxx name represents "figurename.eps"
%\caption{Insert figure caption here}
%\label{fig_sim}
%\end{figure}

\section{Conclusion}
\label{sec:conclusion}

%\appendix{Odds and ends}
%\begin{itemize}
%    \item On the $\Reyn_c(m=1)/\Reyn_c(m=0) %\simeq 1.008$ issue...there's a discussion in  Krueger, Gross, Di Prima (1966), p 523-524, but it references a solution in a conference proceeding that%is difficult to find. I have some memory of it being discussed in a more recent paper, but I can't find it anywhere...
%\end{itemize}
%\enlargethispage{20pt}

%\ethics{Insert ethics statement here if applicable.}

\dataccess{Insert details of how to access any supporting data here.}

\aucontribute{JSO wrote the manuscript, designed the study, wrote the GQL projection operators, and implemented the original Taylor-Couette Dedalus script. MB added the GQL terms to the TC script, developed data analysis techniques, and performed preliminary simulations. Both authors edited the manuscript.}

@@

\funding{JSO was supported by NASA LWS grant number NNX16AC92G and NASA HTMS grant number 80NSSC20K1280. }

>>>>

Here, we consider Taylor-Couette (TC) flows at moderate inner and outer Reynolds numbers, $|\Reyn_{i,o}| < 3500$. This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$. 
One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space. In the linearly stable regions, transient growth has been observed in TC flow\cite{2002PhFl...14.3475H, 2002PhFl...14.1655M}. 
We can always write the Navier-Stokes equations in schematic form as
\begin{equation}
    \label{eqn:NS-schematic}
    \frac{\partial \mathbf{q}}{\partial t} + \Lop \mathbf{q} = \mathcal{N}(\mathbf{q}),
\end{equation}
where $\Lop$ is a linear operator and $\mathcal{N}$ is a non-linear function of the state vector $\mathbf{q} = \left[ \mathbf{u}  p\right]^T$, where $\mathbf{u}$ is the velocity field and $p$ the pressure.
As noted by \cite{2002PhFl...14.3475H}, the linear operator $\Lop$ for TC flow is a function of its control parameters.
That is, its commutator with its adjoint
\begin{equation}
    [\Lop^\dagger, \Lop] = f(\mu, \eta, \Reyn_i, \Gamma)
\end{equation}
can be zero or non-zero. 
This provides a very useful arena to test the validity of GQL in a system that admits a varying degree of non-normality.

As a first experiment to this end, we consider the \emph{spiral turbulence} flow regime. The classic paper by Coles \cite{1965JFM....21..385C} was the first to note a hysteretic transition to spiral turbulence when the cylinders were counter-rotating: an experiment at fixed $\Reyn_o$ will have spiral turbulence at a certain $\Reyn_i$ if $\Reyn_i$ is reached by \emph{decreasing} from a higher value, but will not have spiral turbulence at the same $\Reyn_i$ if it is reached by \emph{increasing} from a lower value. More recent work has identified both supercritical and subcritical paths to spiral turbulence\cite{2009PhRvE..80d6315M}. In that work, the authors follow spiral turbulence along a decreasing $\Reyn_i$ path and show that the state is maintained even below the linear stability boundary when $\Reyn_o = -3000$ but it relaminarizes below the stability boundary when $\Reyn_o = -1200$.
Using \cite{2009PhRvE..80d6315M} as a guide, we use direct numerical simulations (DNS) at $\Reyn_o = -3398$ and decrease $\Reyn_i$ into the subcritical region. We then run a suite of GQL simulations in both the super- and subcritical regions.

@@

We use the pseudo-spectral Dedalus framework (version 2) \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations. We discretize the three velocity components and pressure onto a basis of Chebyshev polynomials in the radial direction, while $\theta$ and $z$ use Fourier bases.
To evolve the solution forward in time, we use the implicit-explicit Runga-Kutta RK443 time stepper; linear terms are implicitly timestepped while non-linear terms are explicit.
Our initial value solver has been extensively tested for TC flow against the linear and non-linear test cases in \cite{1984JFM...146...45M, 1984JFM...146...65M} including linear growth rates, non-linear wave speeds, and torque measurements; our spiral turbulence DNS results also agree with those in \cite{2009PhRvE..80d6315M,2009PhRvE..80f7301D,2011JFM...668..150D}. 
Resolutions and geometric parameters for all runs are given in table~\ref{tab:simulations}.
\begin{table}[!h]
\caption{Simulations}
\label{tab:simulations}
\begin{tabular}{rlllllllll}
\hline
$\Reyn_i$ & $\Reyn_o$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$, $\Lambda_z$ & Notes\\
\hline
%A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
%B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
%C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
%\hline
900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10 & Supercritical SPT\\
700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10, 20, 30 & Subcritical SPT\\
%\hline
%F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
%G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
\end{tabular}
\vspace*{-4pt}
\end{table}%%%End of the table
We have confirmed our simulations are resolved by ensuring that the highest wavenumber components have spectral power at least $10^5$ smaller than the most energetic scale.

Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.

\section{Linear Stability}
\label{sec:linear}
In order to determine the linear stability bound, we find the minimum $\Reyn_i$ for a given $\Reyn_o$ by solving the eigenvalue problem for equations~\ref{eqn:NS}. We note that for all $\Reyn_o$ plotted in figure~\ref{fig:LSB}, most unstable mode is never axisymmetric. At $\Reyn_o=-3398$, where our non-linear calculations are run, the most unstable mode has $m = 5$ at onset, and 
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{../figs/reo_rei_lsb.pdf}
    \caption{A summary of spiral turbulence runs in the $\Reyn_o-\Reyn_i$ plane showing linear stability boundary. Solid circles shows points where we have conducted GQL analysis; crosses represent DNS steps along the path to subcritical behaviour. }
    \label{fig:LSB}
\end{figure}

@@

\section{Results}
\label{sec:nonlinear}
Figure~\ref{fig:urms_tz_rei700} shows a typical DNS snapshot of the rms velocity perturbation $u'_{rms}$ in the middle of the gap for $\Reyn_i=700$. Despite being well below the linear stability threshold $\Reyn_{i,crit} \simeq 778$, the flow morphology is nearly identical to the supercritical case $\Reyn_i = 900$. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../figs/urms_tz_rei_700_reo-3398_000278.png}
    \caption{Slice of fluctuating rms velocity in the middle of the radial domain at $\Reyn_i = 700$, $\Reyn_o = -3398$. This solution is well below the linear stability threshold at }
    \label{fig:urms_tz_rei700}
\end{figure}
\subsection{Supercritical Spiral Turbulence}
We first consider the supercritical case with $\Reyn_i = 900$. We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.  Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$. The first important thing to note is that none of these low-order models are able to maintain the spiral structure. However, even including a single additional low mode, $\Lambda = 1$, shows a significant improvement in flow morphology compared to QL. It is quite interesting to note that QL \emph{does} feature some degree of spatial intermittency--there are clearly separated regions hosting spiral waves interspersed with featureless, laminar patches. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figs/rei900_snapshots.pdf}
    \caption{RMS velocity perturbations in the $z-\theta$ plane at $r=R_{mid}$ for $\Reyn_i=900, \Reyn_0=-3398$. The upper left panel shows DNS, the upper right panel shows QL, and the remaining four are GQL with $\Lambda = 1,3,5,10$. Even at $\Lambda = 10$, spiral turbulence is not apparent.}
    \label{fig:rei900_snapshots}
\end{figure}

Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda$. 
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. On the left,  $\Reyn_i = 900$ is above the linear stability boundary, while the right shows $\Reyn_i = 700$, below the stability boundary. Starting from a saturated spiral turbulence state at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10$. For $\Reyn_i = 700$, we also use $\Lambda = 20, 30$ At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. }
    \label{fig:ke_vs_t_rei900}
\end{figure}
Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating. As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT). Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches is closests approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.

This is notable for a few reasons. At $\Reyn_i=900, \Reyn_o=-3398$ the system is not in a hysteretic region of parameter space. Thus, we do not expect this behaviour to be due to GQL picking up other, coexisting solutions. 
It also highlights the delicate balance an intermittent laminar/turbulent solution represents. The turbulence in this solution has a very broad range of scales, as evidenced by figure~\ref{fig:urms_tz_rei700}. Understanding how it exchanges energy with the mean flow is crucial to understanding both the saturation of the underlying linear instability as well as the maintenance of the non-linear state. 
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/rei900_story.pdf}
    \caption{Kinetic energy (left) and flow morphologies (right) for $\Reyn_i = 900$, $\Lambda = 5$ covering one full oscillation cycle. The panels on the right show three times corresponding to the vertical lines in the kinetic energy plot on the left. At the kinetic energy minimum (point 1), the flow shows a IPS-like morphology, while an INT burst occurs at the knee (point 2), and a more space-filling turbulent state occurs at maximum energy (point 3).}
    \label{fig:rei900_lambda5_story}
\end{figure}

@@

\subsection{Subcritical spiral turbulence}
In order to study subcritical spiral turbulence, we ran a series of DNS with decreasing $\Reyn_i$. Each is initialised with the last timestep of the prior run and run for one viscous time. Once we reached $\Reyn_i = 700$, we run a DNS for a final viscous time and then repeat the procedure from the supercritical GQL study. HThe right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$. Our very preliminary data for $\Lambda = 30$ show that it recovers from the decay and begins to climb in energy again, similar to the early behaviour from the supercritical runs.\footnote{\textbf{Note to reviewers}: this simulation is still running as of the 31 July due date for the initial manuscript. By the time refereeing begins, we will update this plot and section.}

\section{Conclusion}
\label{sec:conclusion}
We have presented a series of simulations demonstrating the performance of GQL at various cutoffs for Taylor-Couette flow in the spiral turbulence regime. GQL is a significant improvement over QL models in representing mean quantities, including the deviation from circular Couette flow. However, even at the highest cutoffs studied here, it does not retain an unambiguous spiral form in either the sub- or supercritical regimes. Prior work has focused on the utility of GQL in presenting a more accurate \emph{statistical} picture of turbulent flows than QL \cite{2016PhRvL.116u4501M, 2017JFM...810..412T,2018RSPSA.47480422T,2019Kellam}. This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realizations \cite{2022arXiv220505513M}.

However, we have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow. In future work, we will consider the ability of GQL to maintain exact coherent states identified by Deguchi et al \cite{2014PhRvL.112r4502D} and the dynamics of which were studied recently by Wang and coworkers \cite{2022arXiv220712990W}. Understanding how these interactions work to shape subcritical TC turbulence represents a promising new avenue to understanding how ECS are maintained and how they in turn sustain turbulence below linear stability bounds. In particular, one can begin to understand the relative importance of

\dataccess{All code for non-linear initial value problems, linear eigenvalue and non-modal stability analyses, and plotting are available at \url{https://github.com/jsoishi/GQL_TC}. Dedalus version 2 is available at \url{https://github.com/DedalusProject/dedalus/tree/v2_master}. eigentools is available at \url{https://github.com/DedalusProject/eigentools}. Simulation data is available upon request.}

\aucontribute{JSO wrote the manuscript, designed the study, wrote the GQL projection operators, and implemented the original Taylor-Couette Dedalus script. MB added the GQL terms to the TC script, developed data analysis techniques, and performed preliminary simulations.}

@@

\funding{JSO was supported by NASA LWS grant number NNX16AC92G and NASA HTMS grant number 80NSSC20K1280. MB was supported in part by a Bates College Travel Grant.}
